---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
    # Process .pdf generated by tikzDevice to .png
    pdf2png <- function(path, options) {
        if (!grepl(".pdf$", path)) return(path)
        outpath <- gsub(".pdf", ".png", path)
        texpath <- gsub(".pdf", ".tex", path)
        img <- magick::image_read(path, density=300)
        magick::image_write(img, outpath, format="png")
        unlink(path)
        unlink(texpath)
        return(outpath)
    }

    knitr::opts_chunk$set(
        collapse = TRUE,
        comment = "#>",
        fig.path = "man/figures/README-",
        fig.width = 3,
        fig.height = 3,
        fig.show = "hold",
        out.width="50%",
        dev = "tikz",
        external = TRUE,
        fig.process = pdf2png
    )

    library(tikzDevice)
    library(ggplot2)
    library(ggtikz)
    options(tikzLatexPackages = c(getOption("tikzLatexPackages"), "\\usetikzlibrary{calc}"))
```

# ggtikz

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/ggtikz)](https://CRAN.R-project.org/package=ggtikz)
[![codecov](https://codecov.io/gh/osthomas/ggtikz/branch/main/graph/badge.svg?token=0LPNGPFO5Z)](https://codecov.io/gh/osthomas/ggtikz)
[![R-CMD-check](https://github.com/osthomas/ggtikz/workflows/R-CMD-check/badge.svg)](https://github.com/osthomas/ggtikz/actions)
<!-- badges: end -->

ggtikz allows you to annotate plots created using
[ggplot2](https://ggplot2.tidyverse.org/)
with arbitrary TikZ code when rendering
them with the [tikzDevice](https://CRAN.R-project.org/package=tikzDevice).
The annotations can be made using data coordinates, or with coordinates relative
to a specified panel or the whole plot.

Plots with multiple panels (via `facet_grid()` or `facet_wrap()`) are supported.

For a few examples, see the [examples vignette](https://github.com/osthomas/ggtikz/blob/main/doc/examples.pdf).

## Installation

You can install the latest ggtikz release from CRAN with:

``` {r, eval = FALSE}
install.packages("ggtikz")
```

Or get the development version from GitHub:

``` {r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("osthomas/ggtikz", ref = "devel")
```

## New in Devel
[![codecov](https://codecov.io/gh/osthomas/ggtikz/branch/devel/graph/badge.svg?token=0LPNGPFO5Z)](https://codecov.io/gh/osthomas/ggtikz)
[![R-CMD-check](https://github.com/osthomas/ggtikz/workflows/R-CMD-check/badge.svg?branch=devel)](https://github.com/osthomas/ggtikz/actions)

* The TikZ coordinates used with transformed scales can now be automatically transformed with the `transform = TRUE` (now the default) option for `ggtikzAnnotation`.

``` {r transform-example}
p_log <- ggplot(mtcars, aes(mpg, disp)) + geom_point() + scale_x_continuous(trans="log10")
canvas_log <- ggtikzCanvas(p_log)

# Untransformed coordinates: wrong position
annot_log <- ggtikzAnnotation(
    "\\fill[red] (1,100) circle (2mm);
    \\node[anchor=west, text=red] at (1, 100)
        {The circle is not at (1,100)!};
    ", xy = "data", transform = FALSE, panelx = 1, panely = 1
)

# Transformed coordinates: correct position
annot_log2 <- ggtikzAnnotation(
    "\\fill[blue] (20,200) circle (2mm);
    \\node[anchor=south, text=blue] at (20, 200)
        {This circle is at ({20,200})!};
    ", xy = "data", transform = TRUE, panelx = 1, panely = 1
)

p_log
canvas_log + annot_log + annot_log2
```

* Plots can be automatically unclipped to accommodate annotations extending beyond the plot boundaries. See `unclip()` and `set_ggtikz_unclip_hook()`.

* `Inf` and `-Inf` in TikZ coordinate specifications can be automatically replaced
by the maximum allowed value for the annotation's reference frame with the
`transform = TRUE` (now the default) option for `ggtikzAnnotation`.


## Basic Usage

1. Create  a ggplot.
2. Add annotations with `ggtikz()`.

```{r basic-usage}
library(ggplot2)
library(ggtikz)

p <- ggplot(mtcars, aes(disp, mpg)) + geom_point() # 1.
## tikz("plot.tikz")
ggtikz(p,
    "\\fill[red] (0.5,0.5) circle (5mm);",
    xy = "panel", panelx = 1, panely = 1)
## dev.off()
## Render with LaTeX ...
```

## Advanced Usage

1. Create a ggplot.
2. Create a ggtikz canvas from the plot with `ggtikzCanvas()`.
3. Create ggtikz annotations with `ggtikzAnnotation()`.
4. Add the annotations to the canvas.
5. Draw the plot and the annotations using tikzDevice.

```{r advanced-usage}
library(ggplot2)
library(ggtikz)

p <- ggplot(mtcars, aes(disp, mpg)) + geom_point()  # 1.
canvas <- ggtikzCanvas(p)                           # 2.
annot1 <- ggtikzAnnotation(                         # 3.
    "
    \\draw (0,0) -- (1,1);
    \\draw (0,1) -- (1,0);
    \\fill[red] (0.5,0.5) circle (5mm);
    ",
    xy = "panel", panelx = 1, panely = 1
)
annot2 <- ggtikzAnnotation(                          # 3.
    "\\draw[<-] (400,20) -- ++(0,3) node[at end, anchor=south] {(400,20)};",
    xy = "data", panelx = 1, panely = 1
)

## tikz("plot.tikz")
p                                                   # 4. + 5.
canvas + annot1 + annot2
## dev.off()
## Render with LaTeX ...

```
